---
title: "Anomaly Detection"
author: "Null Wranglers"
editor: visual
---

## find outliers

histogram of dew percent

```{r}
#| code-fold: true
#| code-summary: "Load Libraries"

suppressWarnings(library(tidyverse))
library(knitr)
library(lubridate)
library(ggplot2)
library(dplyr)
```

```{r}
#| code-fold: true
#| code-summary: "Read in Data"
# read in data
anomaly_data <- read.csv("data/study_data.csv")
head(anomaly_data)
```

```{r}
summary(anomaly_data$Percent.DEW)
```

```{r}
ggplot(anomaly_data, aes(x = Percent.DEW)) +
  geom_histogram(binwidth = 3, fill = "blue", color = "black", alpha = 0.8) +
  labs(title = "Histogram of Percent DEW", x = "Percent DEW", y = "Frequency")
```

```{r}
anomaly_25 <- anomaly_data %>% 
  filter(Percent.DEW>19)

ggplot(anomaly_25, aes(x = Percent.DEW)) +
  geom_histogram(binwidth = 2, fill = "blue", color = "black", alpha = 0.8) +
  labs(title = "Histogram of Percent DEW", x = "Percent DEW", y = "Frequency")
```

whisker and box plot by semester by college

```{r}

```

Summary of quartiles by semester and by college - what is the percentage for outliers - say more than 50% - High Dew

```{r}
ggplot(anomaly_data, aes(y = College, x = Percent.DEW)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  labs(title = "Boxplot of Percent DEW by College",  x = "Percent DEW", y="")

```

pull a df of highest dew and highest pass

count of how from each college in new df scatter plot colored/clustered by college

add a label either high or low (high being over 55% dew percentage) - scatter plot courses and color by label)
