---
title: "Exploratory Data Analysis of Grade Outcomes Data"
author: "The Null wranglers"
editor: visual
---

```{r message=FALSE, output=FALSE}
#| code-fold: true
#| code-summary: "Load Libraries"

suppressWarnings(library(tidyverse))
library(knitr)
library(lubridate)
library(ggplot2)
```

```{r}
#| code-fold: true
#| code-summary: "Read in Data"
# read in data
dew_data <- read.csv("data/clean_dew_data.csv")
kable(head(dew_data))

```

Add brief description of data

## Total number of courses per semester

Use some from the code.qm - just add labels

## Overall Averages

Calc the mean averages overall - high - enrollemnt, pass, d, e, w, dew

Visualiaze in bar

## Averages by College

Calc the mean averages overall - high - enrollemnt, pass, d, e, w, dew

Visualiaze in bar

## Colleges DEW chart and total enrollment

Bar of dew total per college

## find outliers

what is the percentage for outliers - say more than 50% - High Dew

class over 30 enrollment? lower

class under 30 enrollment? upper

per semester?

What is the percentage for outliers - less than 5% - Low Dew
