---
title: "Exploratory Data Analysis of Grade Outcomes Data"
author: "The Null wranglers"
editor: visual
---

```{r message=FALSE, output=FALSE}
#| code-fold: true
#| code-summary: "Load Libraries"

suppressWarnings(library(tidyverse))
library(knitr)
library(lubridate)
library(ggplot2)
```

```{r}
#| code-fold: true
#| code-summary: "Read in Data"
# read in data
dew_data <- read.csv("data/clean_dew_data.csv")
kable(head(dew_data))

```

Add brief description of data

## Total number of courses per semester - K

```{r}
#| code-fold: true
#| code-summary: "summary of total courses"

# add the column names
table1<- dew_data %>% group_by(College) %>% count(TERM_LD)
table1 <- table1 %>% pivot_wider(names_from = TERM_LD, values_from = n)
kable(table1, caption = "Total Course Offering for Each College per Semester"
      )
```

After examining this data, we are limiting the colleges to x,y, z because a, b,c .

```{r}
#| code-fold: true
#| code-summary: "remove low course offering colleges"
dew_data <- dew_data %>% filter(College != "College of Medicine - Tucson",
                                College != "College of Nursing",
                                College != "College of Public Health",
                                College != "Colleges of Letters Arts & Sci",
                                College != "Graduate College",
                                College != "James E Rogers College of Law",
                                College != "R Ken Coit College of Pharmacy",
                                College != "W.A. Franke Honors College")

```

```{r }
#| code-fold: true
#| code-summary: "summary of total courses"

# table2 <- dew_data %>% group_by(College) %>% count(Total.Student.Count)
# kable(table2, caption = "Total Student Enrollment for Courses in Each College"
#       )
```

```{r fig.width=11}
dew_data %>% 
  group_by(TERM_LD, Course.Level) %>% 
  ggplot(aes( x = TERM_LD,
              fill = Course.Level))+
  geom_bar(position = "dodge")+
  geom_text(stat='count', aes(label=..count..), vjust=-1, hjust= 1)
```

Use some from the code.qm - just add labels

## Overall Averages - R

Calc the mean averages overall - high - enrollemnt, pass, d, e, w, dew

Visualiaze in bar

## Averages by College -R

Calc the mean averages overall - high - enrollemnt, pass, d, e, w, dew

Visualiaze in bar

## Colleges DEW chart and total enrollment - table -K

Bar of dew total per college

## find outliers

what is the percentage for outliers - say more than 50% - High Dew

class over 30 enrollment? lower

class under 30 enrollment? upper

per semester?

What is the percentage for outliers - less than 5% - Low Dew
