---
title: "Exploratory Data Analysis of Grade Outcomes Data"
subtitle: "Proposal Final Project"
author: "The Null Wranglers Team"
format: html
editor: visual
---

```{r message=FALSE, output=FALSE}
#| code-fold: true
#| code-summary: "Load Libraries"

suppressWarnings(library(tidyverse))
library(knitr)
library(lubridate)
library(ggplot2)
```

This section covers an Exploratory Data Analysis of the data. All data for this project has been provided by the University of Arizona. This type of analysis is vital in order to gain a full understanding of the data that will guide future analysis techniques. The project will utilize two separate data sets.

# Grade Value Data

The first one contains information about the breakdown of what grade values students received in a a course. A description for the columns of the data can be found in the proposal.

```{r}
#| code-fold: true
#| code-summary: "Read in Data"
# read in data
dew_data <- read.csv("data/clean_dew_data.csv")
kable(head(dew_data))

```

This data set contains 13283 observations with a total of 24 variables.

## Courses Per Semester and Enrollment

The data set spans over 6 different semester that start from the Fall of 2018 and go until the Spring of 2021. During a cleaning process the winter and summer semester were removed. The table below shows the breakdown of course offerings by college.

```{r}
#| code-fold: true
#| code-summary: "summary of total courses"

# add the column names
table1<- dew_data %>% group_by(College) %>% count(TERM_LD)
table1 <- table1 %>% pivot_wider(names_from = TERM_LD, values_from = n)
kable(table1, caption = "Total Course Offering for Each College per Semester")
```

After examining this data, the decision was made to remove colleges that offered less than 50 courses per semester.

```{r}
#| code-fold: true
#| code-summary: "remove low course offering colleges"
dew_data <- dew_data %>% filter(College != "College of Medicine - Tucson",
                                College != "College of Nursing",
                                College != "College of Public Health",
                                College != "Colleges of Letters Arts & Sci",
                                College != "Graduate College",
                                College != "James E Rogers College of Law",
                                College != "R Ken Coit College of Pharmacy",
                                College != "W.A. Franke Honors College")

```

After removing these columns, an examination of the total enrollment in these courses was done.

```{r }
#| code-fold: true
#| code-summary: "summary of total enrollment"
# total enrollment table
enrollment <- aggregate(Total.Student.Count~College + TERM_LD, data=dew_data, sum)
enrollment <- enrollment %>% pivot_wider(names_from = TERM_LD, values_from = Total.Student.Count)
kable(enrollment, caption =  "Total Student Enrollment for Each College per Semester")
```

It was then important to explore the grade value outcomes. In this project, the grade value outcome will be used to determine success and failure for a class. A successful grade outcome will be a letter grade of C or higher. A failure grade will be a letter grade of D, E, or W. An E grade is the representation of an fail grade at the University of Arizona as this institution uses F grade values in Pass/Fail courses only and a W grade represents a withdraw from the course by the student. The data does not contain information for students who drop the course within the allowed drop/add period at the beginning of the semester. The table below shows the D.E.W. grade counts in each college per semester.

```{r}
#| code-fold: true
#| code-summary: "summary of total enrollment"
dew_total <- aggregate(DEW_COUNT~College + TERM_LD, data=dew_data, sum)
dew_total <- dew_total %>% pivot_wider(names_from = TERM_LD, values_from = DEW_COUNT)
kable(dew_total, caption =  "Total D.E.W. Grade Counts for Each College per Semester")
```

After these summaries were, a decision was made to further reduce the data to just the Top 5 Colleges for both enrollment and D.E.W. counts.

```{r}
#| code-fold: true
#| code-summary: "summary of total enrollment"
dew_data <- dew_data %>% filter(College != "College of Applied Sci & Tech",
                                College != "College of Education",
                                College != "College of Engineering",
                                College != "College of Fine Arts")
```

## Overall Averages 

Once the data was reduced, the remaining data contained 8707 observations with 24 variables. Averages for course enrollment and well as D.E.W. averages were calculated and are displayed below.

```{r}
#| code-fold: true
#| code-summary: "average number of students per course"
student_per_course <- aggregate(Total.Student.Count ~College + TERM_LD, data=dew_data, mean)
# round to nearest whole digit
student_per_course$Total.Student.Count <- round(student_per_course$Total.Student.Count)
# pivot wider
student_per_course <- student_per_course %>% pivot_wider(names_from = TERM_LD, values_from = Total.Student.Count)
# display table
kable(student_per_course, caption =  "Average Number of Students per Course by College Each Semester")
```

```{r}
#| code-fold: true
#| code-summary: "average number of students per course"
dew_per_course <- aggregate(Percent.DEW ~College + TERM_LD, data=dew_data, mean)
# round to 2 digits
dew_per_course$Percent.DEW <- round(dew_per_course$Percent.DEW, 2)
# pivot wider
dew_per_course <- dew_per_course %>% pivot_wider(names_from = TERM_LD, values_from = Percent.DEW)
# display table
kable(dew_per_course, caption =  "Average Number of D, F, and W Grade(percentage) per Course by College Each Semester")
```

Despite enrollment staying steady, we see a noticeable decrease in DEW outcomes for the spring of 2020. Followed by an increase for the fall of 2020.

## Averages by Percentage

As certain courses had total enrollment over 1000, the issue that the D.E.W. count from these courses could skew the data was considered. The decision to base further analysis on the D.E.W. percentage was made.

```{r}
# Assuming you have the ggplot2 library installed
# If not, install it using install.packages("ggplot2")
library(ggplot2)

dew_data1 <- read.csv("data/study_data.csv")

# Define the order of variables
variable_order <- c( "Percent.Passed", "Percent.D.Grade", "Percent.E.Grade", "Percent.W.Grade", "Percent.DEW")

# Creating a data frame with the means
means_data <- data.frame(
  Variable = factor(variable_order, levels = variable_order),
  Mean_Value = c(
    mean(dew_data1$Percent.Passed),
    mean(dew_data1$Percent.D.Grade),
    mean(dew_data1$Percent.E.Grade),
    mean(dew_data1$Percent.W.Grade),
    mean(dew_data1$Percent.DEW)
  )
)

# Creating a bar plot
ggplot(means_data, aes(x = Variable, y = Mean_Value, fill = Variable)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(Mean_Value, 2)), vjust = -0.5, size = 3) +  # Add text labels
  labs(title = "Mean of Percentages of Pass and D.E.W Grades",
       x = "Variables",
       y = "Mean Value") +
  theme_minimal()+
  theme_linedraw()+
  theme(legend.position = "bottom", panel.grid = element_blank())
```

```{r}
# Assuming you have the dplyr, knitr, and stringr libraries installed
# If not, install them using install.packages(c("dplyr", "knitr", "stringr"))
library(dplyr)
library(knitr)
library(stringr)

# Group by College and calculate the mean for each grade-related column
college_means_data <- dew_data %>%
  group_by(College) %>%
  summarise(PASS_GRADE_COUNT = mean(PASS_GRADE_COUNT),
            D_GRADE_COUNT = mean(D_GRADE_COUNT),
            FAIL_GRADE_COUNT = mean(FAIL_GRADE_COUNT),
            WITHDRAW_GRADE_COUNT = mean(WITHDRAW_GRADE_COUNT),
            DEW_COUNT = mean(DEW_COUNT))

# Remove underscores from column names
names(college_means_data) <- str_replace_all(names(college_means_data), "_", " ")

# Round off mean values to two decimal points
college_means_data[, -1] <- round(college_means_data[, -1], 2)

# Create a kable table for means of each grade-related column for each college
kable(college_means_data, 
               caption = "Means for Total Pass, D, E, W, and Total DEW Counts Across Colleges")

```

# Course Data

The second data set contain information about the attributes of the course. The first data set was merged with this using the subject code, course code and term offered. A full cleaning script can be found in the extra folder in the GitHub repository. The merged data contains 6568 observations and 70 variables.

```{r}
#| code-fold: true
#| code-summary: "Load data"
# read in data
study_data <- read.csv("data/study_data.csv")
kable(head(study_data))
```

## Time and Day Totals

An exploration of the time of day courses were offered was done by each college. It is important to note that courses with more than one section may be offered at different times and are included this way in this data.

```{r}
#| code-fold: true
#| code-summary: "Total Courses for Time of day"

# Create a table for each time slot
time_slots <- c("Early_Morning", "Mid_Morning", "Early_Afternoon", "Mid_Afternoon", "Evening", "Asynchronous")

tables_list <- lapply(time_slots, function(slot) {
  table_data <- study_data %>%
    group_by(College) %>%
    summarise(Mean_Classes_Attended = round(sum(get(slot), na.rm = TRUE), 2))
  
  names(table_data)[2] <- paste("", slot)
  table_data
})

# Merge tables into a single table by College name
merged_table <- Reduce(function(x, y) merge(x, y, by = "College", all = TRUE), tables_list)

# Display the merged table
kable(merged_table)
```

We then considered the total number of classes offered on each day of the week and asynchronously.

```{r}
#| code-fold: true
#| code-summary: "Total courses for Day of Week"

# Create a table for each day of the week
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday", "Asynchronous")

tables_list <- lapply(weekdays, function(day) {
  table_data <- study_data %>%
    group_by(College) %>%
    summarise(Mean_Classes_Attended = round(sum(get(day), na.rm = TRUE), 2))
  
  names(table_data)[2] <- paste("Mean", day)
  table_data
})

# Merge tables into a single table by College name
merged_table <- Reduce(function(x, y) merge(x, y, by = "College", all = TRUE), tables_list)

# Display the merged table
kable(merged_table)
```

We removed Saturday and Sunday from future analysis.

## Modality

We explored the totals for each mode of delivery for each college and semester.

```{r}
#| code-fold: true
#| code-summary: "Total Courses for Mode"
# Mean Number of students enrolled in in-person vs online vs … (BC to BH) for different colleges - Utkarsha

# Create a table for each course type
course_types <- c("In_Person", "Full_Online", "IntractTV", "Hybrid", "Live_Online")

tables_list <- lapply(course_types, function(course) {
  table_data <- study_data %>%
    group_by(College) %>%
    summarise(Mean_Enrollment = round(sum(get(course), na.rm = TRUE), 2))
  
  names(table_data)[2] <- paste("Enrollment_", course)
  table_data
})

# Merge tables into a single table by College name
merged_table <- Reduce(function(x, y) merge(x, y, by = "College", all = TRUE), tables_list)

# Display the merged table
kable(merged_table)
```

After this we no longer considered the mode of IntractTV.

## Session

We then looked at the total courses offered for each different session length.

```{r}
#| code-fold: true
#| code-summary: "Total Courses for Each Session"

# Create a table for each course type
course_types <- c("Reg_Session", "First_Half_Session", "Second_Half_Session", "First_Third_Session", "Second_Third_Session", "Third_Third_Session", "Ten_Week", "Thirteen_Week", "Other")

tables_list <- lapply(course_types, function(course) {
  table_data <- study_data %>%
    group_by(College) %>%
    summarise(Mean_Enrollment = round(sum(get(course), na.rm = TRUE), 2))
  
  names(table_data)[2] <- paste("", course)
  table_data
})

# Merge tables into a single table by College name
merged_table <- Reduce(function(x, y) merge(x, y, by = "College", all = TRUE), tables_list)

# Display the merged table
kable(merged_table)
```

After this, we decided to only consider Regular Session, First Half Session, and Second Half Session in further analysis.

## Summary

After exploring the data, we were able to remove some colleges and course attributes from further analysis. We were also able to gain a full understanding of the information in the data and begin to develop ideas for further analysis.
